---
title: "HW4"
author: "Qi Qi"
date: "9/25/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 3.3.2

1.

$$f(x;\theta)=\frac{1-cos(x-\theta)}{2\pi}$$
$$\ell=\sum_{i=1}^nlogf(x_i;\theta)=\sum_{i=1}^nlog(1-cos(x_i-\theta))-nlog(2\pi)$$
```{r}
x <- c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
       2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)
logL <- function(theta) {
  sum(log(1 - cos(x - theta))-log(2 * pi))
}
theta <- seq(-pi , pi , by = .01)
logl <- sapply(theta, logL)
plot(theta, logl)

```

2.

$$E(X|\theta)=\int_0^{2\pi}\frac{x(1-\cos(x-\theta))}{2\pi}dx=\int_0^{2\pi}\frac{x}{2\pi}dx-\int_0^{2\pi}\frac{x\cos(x-\theta)}{2\pi}dx$$
$$=\frac{x^2}{4\pi}|_{0}^{2\pi}-\frac{x}{2\pi}\sin(x-\theta)|_0^{2\pi}+\int_0^{2\pi}\frac{\sin(x-\theta)}{2\pi}dx$$
$$=\pi-\sin(-\theta)-\frac{\cos(x-\theta)}{2\pi}|_0^{2\pi}=\pi+\sin\theta$$
$$\Rightarrow \bar{X}_n=\pi+\sin\tilde{\theta}_n \ \Rightarrow \tilde{\theta}_n=\arcsin(\bar{X}_n-\pi)$$

3.

$$\ell'=\sum_{i=1}^n\frac{\sin(\theta-x_i)}{1-\cos(\theta-x_i)}$$
$$\ell''=\sum_{i=1}^n\frac{\cos(\theta-x_i)[1-\cos(\theta-x_i)]-[\sin(\theta-x_i)]^2}{[1-\cos(\theta-x_i)]^2}$$
$$=\sum_{i=1}^n\frac{\cos(\theta-x_i)-1}{[1-\cos(\theta-x_i)]^2}=\sum_{i=1}^n\frac{1}{\cos(\theta-x_i)-1}$$

```{r}
init <- asin(mean(x) - pi)
g <- function(theta){
  sum(sin(theta - x) / (1 - cos(theta - x)))
}
dg <- function(theta){
  sum(1 / (cos(theta - x) - 1))
}
newton<- function(fun, dfun, x0, eps, maxit){
  for (i in 1:maxit) {
    x1 <- x0 - fun(x0) / dfun(x0)
    if (abs(x1 - x0) < eps | abs(fun(x1)) < eps)
      return(x1)
    x0 <- x1
  }
  return(NA)
}
newton(g, dg, init, 1e-6, 1000)

```

The MLE of $\theta$ is 0.0031 with initial value $\tilde{\theta}_n$.

4.

```{r}
newton(g, dg, 2.7, 1e-6, 1000)
newton(g, dg, -2.7, 1e-6, 1000)
```
The MLE for $\theta$ is 2.8484 with initial value $\theta_0=2.7$; MLE for $\theta$ is -2.6689 starting at $\theta_0=-2.7$.

5.

These MLE's could be classified into 197 groups due to very slight differences. So, I round MLE to 6 decimal places. Since there are at most 46 initial values in one group, I report the minimum and maximum of initial values in each group instead of listing all values in each group. 

```{r}
init <- seq(-pi, pi, length.out = 200)
mle <- sapply(init, function(x) newton(g, dg, x, 1e-6, 1000))
apprmle <- round(mle, digits = 6)
tab <- as.data.frame(cbind(init, apprmle))
aggregate(init ~ apprmle, data = tab, FUN = function(x) c(min = min(x), maxi = max(x) ))
```


# Exercise 3.3.3

1.

```{r}
beetles <- data.frame(
    days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
    beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))

growth <- nls(beetles ~ K * beetles[which(days == 0)] / (beetles[which(days == 0)] + (K - beetles[which(days == 0)]) * exp(- r * days)), data = beetles, start = list(r = .15, K = 1000), trace=TRUE)
summary(growth)

```

2.

I set r and K as the values in the trace of previous iteration. The sum squared errors are also obtained from previous iteration.

```{r}

r <- c(0.15, 0.1275481, 0.1197121, 0.1185102, 0.1183094, 0.1182753, 0.1182696, 0.1182686)
K <- c(1000, 1005.5576477, 1039.3710445, 1048.3579593, 1049.2521408, 1049.3815481, 1049.4029151, 1049.4065121)

error <- c(164487.9, 86407.24, 73936.64, 73430.59, 73420.01, 73419.71, 73419.7, 73419.7)

z <-  cbind(r, K, error)
x <-  seq(min(r), max(r), length.out = nrow(z))
y <-  seq(min(K), max(K), length.out = ncol(z))

contour(x, y, z, xlab = "r", ylab = "K")

```

3.

